---
title: "stat_modeling_homework6"
author: "Yujin Jeong"
date: "2019년 4월 21일"
output: word_document
---

# 1 

## (a)

```{r}
t<-c(1:11)
y<-c(1,1,2,3,4,6,8,13,21,27,45)
plot(y~t)
```

I can see the curve in the picture. The curve looks like a growth curve model because it is concave.

## (b)

y=B0*B1^t*e
ln(y)=B0+tln(B1)+e
B0 and ln(B1) are the parameters and they are just constant numbers.

## (c)

```{r}
lny<-log(y)
plot(lny~t)
```

Yes

## (d)

```{r}
model<-lm(lny~t)
summary(model)
```

## (e)

```{r}
lny12<--0.54334+t*0.38997
lny12

new.data.frame<-data.frame(t=12)
predict(model,newdata=new.data.frame,interval="prediction",level=0.95)
```

## (f)

```{r}
y1<-y[1:length(y)-1]
model1<-lm(y[2:length(y)]~y1+0)
summary(model1)
```

The point estimate is 1.53197. It means that the number of diseases of that year is about 1.53 times bigger than the last year's number of diseases.

```{r}
confint(model1,'y1',level=0.95)
```

(1.401519, 1.662426)
We are 95% confident that the number of diseases of that year is bigger than between 1.401519 and 1.662426 times than the last year's number of diseases.

## (g)

```{r}
y12<-exp(4.13629)
y12
```


# 2

## (a)

```{r}
year<-c(1,1,1,1,2,2,2,2,3,3,3,3)
quarter<-factor(c(1,2,3,4,1,2,3,4,1,2,3,4))
sales<-c(2915,8032,10411,2427,4381,9138,11386,3382,5105,9894,12300,4013)
period<-seq(1:12)
time.sale<-ts(sales,frequency=4,start=1)
sale.dat<-data.frame(sales,period,quarter)

plot(time.sale)
plot(decompose(time.sale))
```

## (b)

We can oberserve that it has linear trend and seasonal effect.

## (c)

```{r}
sale.mod<-lm(sales~period+relevel(quarter,4),data=sale.dat)
summary(sale.mod)
```

## (d)

### i

```{r}
summary(sale.mod)
```

Every independent variables in the model seem important. 

### ii and iii

```{r}
new.data.frame1<-data.frame(period=c(13,14,15,16),quarter=factor(c(1,2,3,4)))
predict(sale.mod,newdata=new.data.frame1,interval="prediction",level=0.95)
```

13247.417 is the point estimate. Prediction interval is (12677.550, 13817.283).

# 3

## (a)

```{r}
x3<-c(8,10,12,14,16,18,20)
y3<-c(6.47,6.17,7.4,10.57,11.93,10.3,14.67)
plot(y3~x3)
```

### i

```{r}
model.test1<-lm(log(y3)~x3)
shapiro.test(resid(model.test1))
library(lmtest)
bptest(model.test1,studentize = F)
```

### ii

```{r}
model.test2<-lm(I(1/y3)~I(1/x3))
shapiro.test(resid(model.test2))
bptest(model.test2,studentize = F)
```

### iii 

```{r}
model.test3<-lm(y3~I(1/x3))
shapiro.test(resid(model.test3))
bptest(model.test3,studentize = F)
```

These three models satisfies the assumptions.

## (b)

I will use model1 and model3.

```{r}
new.data.frame2<-data.frame(x3=16)
predict(model.test1,newdata=new.data.frame2,interval="prediction",level=0.95)
exp(1.99038)
exp(2.733324)
```

```{r}
predict(model.test3,newdata=new.data.frame2,interval="prediction",level=0.95)
```

Model 1 is more precise because it has shorter interval. 
