---
title: "statistical modeling hw2"
author: "Yujin Jeong E00628980"
output: word_document
---

# 1


## (a)


```{r}
y<-c(566.52,696.82,1033.15,1603.62,1611.37,1613.27,1854.17,2160.55,2305.58,3503.93,3571.89,3741.40,4026.52)
x<-c(472.92,1339.75,620.25,568.33,1497.6,1365.83,1687,1639.92,2872.33,3655.08,2912,3921,3865.67)
plot(y~x)
```

## (b)


```{r}
lin<-lm(y~x)
lin
```

y=0.8705x+407.0315

## (c)

b0 : On average, for a hospital with monthly occupied beds of 0, the predicted monthly labor hours required are 407.0315.

b1 : On average, for every one unit increase in monthly occupied beds, the monthly labor hours required increases by 0.8705.

## (d)


```{r}
plot(y~x)
abline(lm(y~x),col="red")
```

yes.

## (e)


```{r}
summary(lin)
qt(0.975,length(y)-2)
```

T statistic is 8.344 which is much greater than 2.201(critical value). Therefore we reject the null hypothesis that there is no significant linear relationship between monthly labor hours required and monthly occupied beds for U.S. Navy Hospitals. There is a significant linear relationship.

## (f)


```{r}
confint(lin, level=0.99)  
```

Interval : (0.5465133,1.194547)

With 99% confidence, for every one monthly occupied bed increase in average,the monthly labor hours required increases between 0.5465133 and 1.194547 on average.

## (g)


```{r}
MSR<-sum((0.8705*x+407.0315-mean(y))^2)
SSE<-sum((y-(0.8705*x+407.0315))^2)
MSE<-SSE/(length(y)-2)
F_stat<-MSR/MSE
F_stat
```

## (h)


```{r}
qf(0.95,1,11)
```

F is 15.59786 which is much greater than 4.84(critical value). Therefore We reject the null hypothesis that there is no significant linear relationship. There is significant linear relationship between monthly labor hours required and monthly occupied beds for U.S. Navy Hospitals.

## (i)


```{r}
t_stat<-8.344
t_stat^2
F_stat
```

They are almost same.


```{r}
F_value<-qf(0.95,1,11)
T_value<-qt(0.975,11)
F_value
T_value^2
```

They are same.

## (j)

the result from (i) says that parts (e) and (h) are testing the same test.

## (k)


```{r}
T_value90<-pt(0.95,11)
c00<-(1/length(y))+(mean(x))^2/sum((x-mean(x))^2)
s<-sqrt(SSE/(length(y)-2))
407.0315-T_value90*sqrt(c00)*s
407.0315+T_value90*sqrt(c00)*s
```

(204.932,609.131)

## (l)


```{r}
y0_hat<-0.8705*3000+407.0315
h00<-1/length(y)+(3000-mean(x))^2/sum((x-mean(x))^2)
y0_hat-T_value*s*sqrt(h00)
y0_hat+T_value*s*sqrt(h00)
```

(2662.4,3374.663)

We are 95% confident that the monthly labor hours required for a US Navy hospitals with 3000 monthly occupied beds is between 2662.4 and 3374.663

## (m)


```{r}
y0_hat-T_value*s*sqrt(h00+1)
y0_hat+T_value*s*sqrt(h00+1)
```

(1953.864,4083.199)

We are 95% confident that the mean monthly labor hours required for all US Navy hospitals with 3000 monthly occupied beds is btween 1953.864 and 4083.199

## (n)


```{r}
SST<-sum((y-mean(y))^2)
SSR<-sum((0.8705*x+407.0315-mean(y))^2)
r<-sqrt(SSR/SST)

SST
SSR
SSE
r^2
```

If r^2 is close to 1, it means that the correlation between independent variables and dependent variables is high. Therefore the monthly labor hours required and monthly occupied beds are highly correlated because 0.863511 is high enough.



# 2

```{r}
y2<-c(21.74,25.41,25.44,25.4,23.91,27.05,26.63,26.98,28.26,27.87,28.4,30.16,32.61,30.33,30.18,33.33)
t<-c(1:16)
```

## (a)


```{r}
plot(y2~t)
```

yes

## (b)


```{r}
lin2<-lm(y2~t)
lin2
```

y=0.603t+22.606

## (c)


```{r}
new_t<-17
new_y<-0.603*new_t+22.606
new_y

new_t<-18
new_y2<-0.603*new_t+22.606
new_y2
```

The prediction of personal coumputer sales in time periods 17 is 32.857. That of personal coumputer sales in time periods 18 is 33.46.

## (d)


```{r}
b1<-0.603
b0<-22.606
SSE2<-sum((y2-(b0+b1*t))^2)
s2<-SSE2/(length(y2)-2)
s<-sqrt(s2)

SSE2
s2
s
```

## (e)


```{r}
summary(lin2)
```

The p-value of the test of b1=0 is 2.093-07. It is small enough to reject the null hypothesis. Therefore we are 99% confident that the trend line relating sales y to time t is significant.

## (f)


```{r}
newdata<-data.frame(t=17)
predict(lin2,newdata,interval="prediction",level=0.95)
```

(29.98197,37.73153)



# 3

```{r}
hotel<-matrix(c(
385, 2, 59, 56.0,
385, 1, 62, 38.0,
1256, 2, 139, 56.0,
1256, 2, 51, 49.0,
1256, 3, 150, 67,
1256, 3, 151, 60.5,
350, 3, 89, 50.5,
292, 2, 44, 49,
185, 3, 101, 55,
290, 2, 101, 65,
3645, 3, 317, 94,
3645, 3, 159, 57,
3645, 3, 257, 67,
820, 2, 98, 66,
360, 2, 116, 54,
200, 2, 130, 63,
131, 2, 32, 46.5,
515, 3, 128, 59,
550, 3, 118, 65,
260, 3, 117, 57,
650, 2, 53, 53,
580, 2, 147, 44,
363, 2, 61, 42,
250, 2, 170, 47,
462, 3, 117, 66,
2140, 2, 190, 43,
2140, 1, 49, 51,
5750, 3, 175, 65,
5750, 2, 169, 41,
5840, 3, 160, 78,
5840, 3, 201, 83.5,
5840, 3, 117, 71,
5840, 2, 148, 46,
3954, 3, 250, 94,
3954, 3, 246, 78),ncol=4,byrow=T)

hotel<-data.frame(hotel)
names(hotel)<-c("x1","x2","x3","y3")

x1<-hotel[,1]
x2<-hotel[,2]
x3<-hotel[,3]
y3<-hotel[,4]
```

## (a)

### (x1)


```{r}
plot(y3~x1)
```

### (x2)


```{r}
plot(y3~x2)
```
### (x3)


```{r}
plot(y3~x3)
```

I am not sure about the model only considering these pictures.

## (b)

### (x1)


```{r}
ln1<-lm(y3~x1)
ln1

SSR1<-sum((0.00286*x1+53.61916-mean(y3))^2)
SST1<-sum((y3-(0.00286*x1+53.61916))^2)
coef1<-SSR1/SST1
coef1
```

### (x2)


```{r}
ln2<-lm(y3~x2)
ln2

SSR2<-sum((14.98*x2+22.96-mean(y3))^2)
SST2<-sum((y3-(14.98*x2+22.96))^2)
coef2<-SSR2/SST2
coef2
```

### (x3)


```{r}
ln3<-lm(y3~x3)
ln3

SSR3<-sum((0.1366*x3+41.1123-mean(y3))^2)
SST3<-sum((y3-(0.1366*x3+41.1123))^2)
coef3<-SSR3/SST3
coef3
```

Second model that independent variable is x2 is the highest proportion of the total variation in the observed room rates.

## (c)


```{r}
qt(.075,length(y3)-2)
```

### (x1)


```{r}
summary(ln1)
```

Its t value is 2.744 which is a little bigger than 2.034515(95% critical value). Therefore we reject the null hypothesis(b1=0). There is the significance of the relationship for each of the models.

### (x2)


```{r}
summary(ln2)
```

Its t value is 4.896 which is bigger than 2.034515(95% critical value). Therefore we reject the null hypothesis(b1=0). There is the significance of the relationship for each of the models.

### (x3)


```{r}
summary(ln3)
```

Its t value is 4.887 which is bigger than 2.034515(95% critical value). Therefore we reject the null hypothesis(b1=0). There is the significance of the relationship for each of the models.

## (d)

I think second model whose independent variable is rating of hotel is the best model. Because its coefficient of determination is highest and there is a huge significance of the relationship .

## (e)


```{r}
newdata1=data.frame(x2=3)
predict(ln2,newdata1,interval="prediction",level=0.95)   
```

(45.2334,90.57569)

## (f)


```{r}
predict(ln2,newdata1,interval="confidence",level=0.95)  
```

(62.74846, 73.06063)